---
title: "Repro_alloc_1"
author: "R.Ward"
date: "3/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


The goal of this preliminary analysis is to look at the allocation to reproduction among BCI's most common species, and compare allocation fractions between PFTs.     

We're interested in:   

repro_output / leaf_litter_output  

seed_ouput / repro_output  

Our PFTs are: drought tolerant and intolerant (dt/di) and early and late successional(es/ls).    
- Adam HB has filtered BCI species list for those with greater than 50 reproductive individuals, giving us a list of 26 species.     
- Joe Wright has extracted the summed reproductive output for these species from two sites, with 62 and 60 traps respectively.  

```{r}
library(tidyverse)
library(readr)
library(skimr)
```

1.  Importing species list:

```{r}
#this file contains the spp list and pfts for the 26 spp identified with more than 50 repro individuals
species_list_26 <- read_csv("species_list_for_repro_alloc_26_mostabund_sp.csv", col_types = cols(
  sp = col_character(),
  Latin = col_character(),
  dmax = col_double(),
  wsg = col_double(),
  e_vs_l = col_factor(),
  dpft = col_factor(),
  repro_abund_2010 = col_double(),
  pft = col_factor()
))
head(species_list_26)

```

2. Importing BCI trait data:  
- Reproductive trait data is from BCITRAITS 
- Units are in g

```{r}
bci_repro_traits <- read_csv("BCITRAITS_20101220 (1).csv") %>%
                    mutate(sp = tolower(`SP$`)) %>%
                    select(sp, FRUIT_DRY, N_SEEDFULL, SEED_DRY, DSPR_DRY)

#view(BCITRAITS_20101220_1_)

head(bci_repro_traits) 
```


3. Importing basal area and reproductive basal area data for the 50 ha plot:
 - provided by Adam HB, units in m2, for year 2010

```{r}

b_area <- read_csv("basal_area_2010.csv") %>%
              select(sp, ba)


rb_area <- read_csv("repro_basal_area_df.csv") %>%
           select(sp, r_ba)

```


4. Joining the sp. list, bci traits, and basal area data:
- checking missing values for our list of 26 species

```{r}
sp_traits <- species_list_26 %>%
                    left_join(bci_repro_traits, by = "sp") %>%
                    left_join(b_area, by = "sp") %>%
                    left_join(rb_area, by = "sp")

head(sp_traits)
skim(sp_traits)

#From the skim, note there are 5 sp for which BCI traits is missing trait values


missing_values <- sp_traits %>%
                  filter(is.na(DSPR_DRY) | is.na(FRUIT_DRY) | is.na(N_SEEDFULL) | is.na(SEED_DRY))

missing_values
```


5. Importing 50-ha plot data  
- This file sums 62 x 0.5m2 traps in the 50 ha plot for which masses were recorded from 25 June 2013 through 29 January 2019
- The types are leaves, flowers, immature fruit and mature fruit (including seeds, entire fruit, capsules, and fruit damaged by vertebrates or insects)
```{r}
mass_data_50haplot <- read_csv("Hanbury_Brown_50ha_20190221.csv", col_types = cols(sp = col_character(), 
                                                                                   type = col_factor(), 
                                                                                   mass = col_double(),
                                                                                   Freq = col_double())
                                                                                      )
skim(mass_data_50haplot)
head(mass_data_50haplot)
```

6. Scaling: 
Reproductive mass:
- Our mass value is a sum of reproductive output from 25 June 2013 through 29 January 2019 (67 months)
- Reproductive output per year =  $R_{mass} / 67_{mo} * 12_{mo} / year = R_{mass} / year$
- Mass per m2 for 62 0.5m2 traps = $R_{mass}/ 31 m^2$
- Reproductive mass per m2 scaled to the 50 ha plot by sp reproductive basal area = $R_{mass} / m^2 * 500,000 m^2 * 1/ BA_{R} (m^2)$

Leaf litter mass:   
- Leaf litter output per year per year = $L_{mass} / 67_{mo} * 12_{mo} / year = L_{mass} / year$  
- Mass per m2 for 62 0.5m2 traps = $L_{mass}/ 31 m^2$  
- Leaf litter mass per m2 scaled to the 50 ha plot by sp total basal area = $L_{mass} / m^2 * 500,000 m^2 * 1/ BA m^2)$ 


```{r}
repro_total_df <- mass_data_50haplot %>%
                  left_join(sp_traits, by = "sp") %>%
                  filter(type != "leaf") %>%
                  group_by(sp)%>%
                  mutate(repro_mass_total = sum(mass)) %>%
                  mutate(repro_mass_per_yr = repro_mass_total * 12 / 67) %>%
                  mutate(repro_mass_per_yr_m2 = repro_mass_per_yr / 31) %>%  
                  mutate(repro_mass_per_yr_indv = repro_mass_per_yr / repro_abund_2010) %>%
                  mutate(repro_output_50_ha = (repro_mass_per_yr_m2 * 500000 / r_ba)) %>%
                  filter(type == "mature fruit") %>%
                  mutate(mature_frt_mass = sum(mass)) %>%
                  select(sp, mature_frt_mass, repro_mass_total, repro_mass_per_yr_indv, repro_mass_per_yr, repro_mass_per_yr_m2, repro_output_50_ha) %>%
                  distinct(sp, .keep_all = T)

repro_total_df

leaf_total_df <- mass_data_50haplot %>%
                 left_join(sp_traits, by = "sp") %>%
                 filter(type == "leaf") %>%
                 group_by(sp) %>%
                 mutate(leaf_mass_per_yr = mass * 12 / 67) %>%
                 mutate(leaf_mass_per_yr_m2 = leaf_mass_per_yr / 31) %>%
                 mutate(leaf_output_50_ha = (leaf_mass_per_yr_m2 * 500000 / ba)) %>%
                 select(sp, leaf_mass_per_yr, leaf_mass_per_yr_m2, leaf_output_50_ha) %>%
                 distinct(sp, .keep_all = T)

leaf_total_df

```


7. Joining all sp specific trait data: 
```{r}
    
Repro_trait_data <- sp_traits %>%
                    left_join(repro_total_df, by = "sp") %>%
                    left_join(leaf_total_df, by = "sp") %>%
                    mutate(frac_repro_output = repro_output_50_ha / leaf_output_50_ha)

head(Repro_trait_data)
skim(Repro_trait_data)

outliers<- Repro_trait_data %>%
   filter(frac_repro_output >= 1.0) %>%
   select(sp, repro_mass_per_yr_m2, ba, r_ba, repro_output_50_ha, leaf_output_50_ha, frac_repro_output)

outliers

#the sp quaras seems to have an error in basal area or reproductive basal area measure
#basal area is reported as 120.8 m2, reproductive basal area is reported as 0.91 m2
#Quaras repro_output per year per m2 = 27.22g, it seems unlikely repro basal area is only .9
#the resulting fraction of repro_output / leaf_output = 87
#removed quaras from below boxplots

Repro_trait_data %>%
        group_by(pft) %>%
        filter(sp != "quaras") %>%
        ggplot(aes(pft, frac_repro_output)) + geom_boxplot()


#arbitrarily expluding potential outliers, where frac_repro_output was greater than 1.5
#even exluding these, the average frac of repro_output/leaf_litter_output >> 0.1

Repro_trait_data %>%
        group_by(pft) %>%
        filter(frac_repro_output < 1.5) %>%
        ggplot(aes(pft, frac_repro_output)) + geom_boxplot()

Repro_trait_data %>%
        group_by(e_vs_l) %>%
        filter(sp != "quaras") %>%
        ggplot(aes(e_vs_l, frac_repro_output)) + geom_boxplot()

Repro_trait_data %>%
        group_by(e_vs_l) %>%
        filter(frac_repro_output < 1.5) %>%
        ggplot(aes(e_vs_l, frac_repro_output)) + geom_boxplot()

Repro_trait_data %>%
        group_by(dpft) %>%
        filter(sp != "quaras") %>%
        ggplot(aes(dpft, frac_repro_output)) + geom_boxplot()


```

8. Calculating expected seed mass, and fraction of reproductive output that is seed:

- Mass in the mature fruit category is treated as fruit (although we are aware that this category includes seeds, entire fruit, capsules, and fruit damaged by vertebrates or insects)
- Expected seed mass is calculatec from mature fruit mass using BCI trait measures for average weight of dried fruit, average number of seeds per fruit, and average weight of dried seeds
- Expected seed mass = Mature_frt_mass / fruit_dry_mass * number_seeds * seed_weight
- Units for the numerator and denominator of this fraction are in g, both are summed over 67 months for 31 m2

```{r}
repro_seeds_df <- Repro_trait_data %>%
                  mutate(expd_seed_mass = ((mature_frt_mass/ FRUIT_DRY) * N_SEEDFULL * SEED_DRY)) %>%
                  mutate(fract_seed_of_repro = expd_seed_mass / repro_mass_total)

head(repro_seeds_df)
skim(repro_seeds_df)

#graphing pft
repro_seeds_df %>%
        group_by(pft) %>%
        ggplot(aes(pft, fract_seed_of_repro)) + geom_boxplot()

#summarizing by pft
repro_seeds_df %>%
        group_by(pft) %>%
        summarize(per_pft_seed_frac_repro = mean(fract_seed_of_repro, na.rm = T), 
                  se = sd(fract_seed_of_repro, na.rm = T)/length(fract_seed_of_repro))
   
#graphing e_vs_l
repro_seeds_df %>%
        group_by(e_vs_l) %>%
        ggplot(aes(e_vs_l, fract_seed_of_repro)) + geom_boxplot()

#graphing dt vs di
repro_seeds_df %>%
        group_by(dpft) %>%
        ggplot(aes(dpft, fract_seed_of_repro)) + geom_boxplot()
```

Next steps:     
1. Find/estimate missing values for bci traits (perhaps from similar sp, if not availibe from J. Wright)  
2. Add additional sp. if possible (esp for underrepresended PFTs in this 26 sp sample)  
3. Correct biases:  
- losses to herbivory (ask J Wright for guidance here)  
- sp biased with respect to spatial sampling  
